|

### 2023-03-12 paper_comparison/appendix.Rmd
update on paper and appendix

---
title: "appendix"
output: pdf_document
date: "2023-01-22"
---

@@

@@

```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value)) %>%
  left_join(
vi_survey_upper %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_upper = first(p_value))) %>%
  left_join(
vi_survey_lower %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_lower = first(p_value))) %>%
  pivot_longer(p_value_upper:p_value_lower) %>%
  mutate(name = ifelse(name == "p_value_upper", "+25%", "-25%")) %>%
  ggplot() +
  geom_point(aes(p_value, value), alpha = 0.2) +
  geom_abline() +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~name) +
  ylab("Adjusted p-value") +
  xlab("p-value")

```

@@

>>>>

---
title: "Supplmentary material - A plot is worth a thousand tests: assessing residual diagnostics with visual inference"
output: pdf_document
author: ""
---

@@

(use LV plot `lvplot` `geom_lv`)

@@

```{r sensitive, fig.cap = "\\label{fig:sensitive}Change of $p$-values with $\\hat{\\alpha}$ inflated by 25% and deflated by 25%."}
vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value = first(p_value)) %>%
  left_join(
vi_survey_upper %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_upper = first(p_value))) %>%
  left_join(
vi_survey_lower %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_lower = first(p_value))) %>%
  pivot_longer(p_value_upper:p_value_lower) %>%
  mutate(name = ifelse(name == "p_value_upper", "+25%", "-25%")) %>%
  ggplot() +
  geom_point(aes(p_value, value), alpha = 0.2) +
  geom_abline() +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~name) +
  ylab("Adjusted p-value") +
  xlab("p-value") +
  coord_fixed()
```

@@

## Curious effect of predictor distribution on conventional test power

### 2023-03-12 paper_comparison/paper_comparison/paper_comparison.Rmd
update on paper and appendix

Effect size in statistics measures the strength of the signal relative to the noise. It is surprisingly difficult to quantify in general, even for simulated data as used in this experiment.

For the non-linearity model, the key items defining effect size are sample size ($n$) and variance of the error term ($\sigma^2$), and so effect size would be roughly calculated as $\frac{\sqrt(n)}{\sigma}$. As sample size increases the effect size would increase, but as variance increases the effect size decreases. However, it is not clear how the additional parameter for the model polynomial order, $k$, should be incorporated. Intuitively, the large $k$ means more complex pattern, which likely means effect size would decrease. For the purposes of our calculations we have chosen to use an approach based on Kullback-Leibler divergence [@kullback1951information], coupled with simulation. This formulation defines effect size to be:

$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$
\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,
$\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients, and $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true.

In the heteroskedasticity model, the key elements for measuring effect size are sample size, $n$, and the ratio of the biggest variance to smallest variance, $b$. Larger values of both would produce higher effect size. However, it is not clear how to incorporate the additional shape parameter, $a$. Thus the same approach is used here, where the formula can be written as:

$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
\noindent where $\boldsymbol{V}$ is the actual covariance matrix of the error term.

Derivations for these equations are provided in the Appendix.

To compute the effect size for each lineup we simulate ??

>>>>

Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample. Importantly, it needs to reflect the treatment we try to measure. According to the underlying simulated model, we define the effect size

$$E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)$$

for the non-linearity model, and

$$E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)$$
for the heteroskedasticity model, where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix, $\sigma^2\boldsymbol{I}$ is the assumed covariance matrix of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual covariance matrix of the error term, $\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a$ is the residual operator, $\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'$ is the hat matrix, $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of predictors including the intercept used in the regression equation, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.

Both equations above are derived from the Kullback-Leibler divergence [@kullback1951information] and the details are provided in the Appendix.
